### 2022-11-04 NSF2022/Description/backgroundv2.tex
started to work on BIE section

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent {\bf MAKE MORE MATHEMATICAL}} \\
{\em Rolf Ryham (lead PI, Fordham University),
Bryan Quaife (PI, Florida State University), and
Yuan-Nan Young (PI, New Jersey Institute of Technology)}

@@

Recently, we have developed a new mathematical model of self-organization.
The model takes ideas from the phase field method 
and leverages boundary integral equations to speed up computations.
To formulate the problem, consider a collection of $N_b$-many rigid bodies
$U_i$ immersed in an incompressible, zero-Reynolds number fluid
$\Omega_t \mathbb{R}^n \setminus \cup_{i=1}^{N_b} U_i$.
The subscript $t$ reinforces that $\Omega_t$ is time-varying. 
We use the grain shape and boundary conditions to capture the local 
properties of the suspended media.
Unlike particles, the bodies, which we will call grains (``using `grains' as an easy find-replace for
when we settle on vocabulary), have finite size.

The first key idea is
to define the grain interactions using a boundary value problem
\begin{align}
  -\mu\Delta \uu + \nabla p &= \mathbf{0},\\
  \nabla \cdot \uu &= 0,\\
-\rho^2 \Delta \eta + f'(\eta) &= 0, \quad \text{ in } \Omega_t,
\end{align}
with boundary conditions
\begin{align}
  \frac{d\xx}{dt} &= \uu,\\
  \uu &= \vv_i + \omega_i \times (\xx - \aa_i)\\
  \eta &= h_i(\xx,t),\quad \text{ for } \xx \in \Gamma_i
\end{align}
and
\begin{equation}
\int_{\Gamma_i} \mu(\nabla \uu + \nabla \uu^T + pI)\nu \,\dif s = 
\int_{\Gamma_i}
\underbrace{(-\rho^{-1} f'(\eta)
  + \rho(\tfrac{1}{2}|\nabla \eta|^2 - \nabla \eta \nabla \eta^T)}_{T_i} \nu \, \dif s
\end{equation}
This is a semi-linear system for evolution of the grains.  At the top,
the Stokes equations describe the velocity $\uu$ and pressure $p$ of the
aqueous phase. In third equation describes the phase transitions of the
scalar order parameter $\eta$ with a decay length $\rho.$
Mathematically, this order parameter is responsible for the attraction
and adhesion between grains that want to align boundaries of like phase.  
Physically, water molecules are polar and $\eta$ describes the orientational
order of water e.g. $\eta = 1$ when water molecules share alignment
and $\eta = 0$ when the water orientations are disordered.
The well-potential $f$ accounts for possibly distinct equilibrium phases,
as occurs with 3-bond and 4-bond hydrogen bonding.

@@

The boundary integral equation (BIE) method is
the idea numerical method for the present study
in terms of computational cost and accuracy.
Boundary integral equations
require linear PDE and so we take $f(\eta) = \tfrac{1}{2}\eta^2$,
yielding in () the so called screened Laplace equation
$-\rho^2 \Delta \eta + \eta$.  This is a Helmholtz equation with purely imaginary
wave number, enabling us to represent the solution
in terms of layer potentials:
\begin{equation}
  \eta(\xx) = \int_{\partial \Omega} \frac{\partial G}{\partial \nu}
  \left(\frac{\xx-\yy}{\rho}\right) \sigma(\yy) \,ds_{\yy}
\end{equation}
where $\frac{\partial G}{\partial \nu}$ is the normal derivative of the
fundamental solution of the screened Laplace equation
(a Bessel function in two space dimensions and
an exponential times the standard fundamental solution of the Laplace equation
in three space dimensions).
This representation allows us to transfer the boundary value problem ()-()
to an equation only involving in terms of domain boundary.

The numerical advantages of doing so are as follows.  If $h = N^{-1}$
is the characeristic grid spacing of the grain boundaries, then a
finite element or finite difference discretization of ()-()
requires on the order of $N_b N^n$ grid points
for dense suspensions.  In contrast BIE require onely 
With finite elements, mesh regeneration
would be needed at each time step since the domain is moving.  
In the boundary integral approach, collocation points can be selected
for the initial grain configuration and then used throughout the
time-course.
As a downside, integral equations require specialized quadrature
routines to resolve self interactions and nearly touching
grain boundaries.  
Finally, the far field boundary conditions which are important for
imposed flow problems automatically hold, whereas with 
finite element and finite difference methods
artificial boundary conditions must be imposed.
\begin{figure}
A figure comparing FE versus BIE meshes
\end{figure}

@@

\subsection{Boundary Integral formulation}
\textbf{Take from prior years proposal}

>>>>

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent {\bf MAKE MORE MATHEMATICAL}} \\
{\em Rolf Ryham (PI, Fordham University),
Bryan Quaife (lead PI, Florida State University), and
Yuan-Nan Young (PI, New Jersey Institute of Technology)}

@@

\subsection{Problem formulation}
Recently, we have developed a new mathematical model of self-organization.
The model takes ideas from the phase field method 
and leverages boundary integral equations to speed up computations.
To formulate the problem, consider a collection of $N_b$-many rigid bodies
$U_i$ immersed in an incompressible, zero-Reynolds number fluid
$\Omega_t \mathbb{R}^n \setminus \cup_{i=1}^{N_b} U_i$.
The subscript $t$ reinforces that $\Omega_t$ is time-varying. 
We use the grain shape and boundary conditions to capture the local 
properties of the suspended media.
Unlike particles, the bodies, which we will call grains (``using `grains' as an easy find-replace for
when we settle on vocabulary), have finite size.

The first key idea is to define the grain interactions using a boundary
value problem
\begin{alignat}{3}
  -\mu\Delta \uu + \nabla p &= \mathbf{0}, \:
  \nabla \cdot \uu = 0, \quad &&\xx \in \Omega, \label {eqn:stokes} \\
  -\rho^2 \Delta \eta + f'(\eta) &= 0, &&\xx \in \Omega,
  \label{eqn:phase}
\end{alignat}
with boundary conditions
\begin{alignat}{3}
  \frac{d\xx}{dt} &= \uu, &&\xx \in \bd\Omega \\
  \uu &= \vv_i + \omega_i \times (\xx - \aa_i), \quad &&\xx \in
  \Gamma_i,\, i=1,\ldots,N_b, \\
  \eta &= h_i(\xx,t),  &&\xx \in \Gamma_i,\, i=1,\ldots,N_b,
\end{alignat}
and
\begin{equation}
\int_{\Gamma_i} \mu(\nabla \uu + \nabla \uu^T + pI)\nu \,\dif s = 
\int_{\Gamma_i}
\underbrace{(-\rho^{-1} f'(\eta)
  + \rho(\tfrac{1}{2}|\nabla \eta|^2 - \nabla \eta \nabla \eta^T)}_{T_i} \nu \, \dif s
\end{equation}
This is a semi-linear system for evolution of the grains.  At the top,
the Stokes equations describe the velocity $\uu$ and pressure $p$ of the
aqueous phase. In third equation describes the phase transitions of the
scalar order parameter $\eta$ with a decay length $\rho.$
Mathematically, this order parameter is responsible for the attraction
and adhesion between grains that want to align boundaries of like phase.  
Physically, water molecules are polar and $\eta$ describes the orientational
order of water e.g. $\eta = 1$ when water molecules share alignment
and $\eta = 0$ when the water orientations are disordered.
The well-potential $f$ accounts for possibly distinct equilibrium phases,
as occurs with 3-bond and 4-bond hydrogen bonding.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral formulation}
Boundary integral equations (BIEs) are a reformulation of elliptic PDEs.
For example, solutions of the Stokes equations~\eqref{eqn:stokes} can be
written as the convolution 
\begin{align*}
  \uu(\xx) = \int_{\partial\Omega} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
\end{align*}
where $G(\xx - \yy)$ is a carefully chosen derivative of the PDE's
fundamental solution and $\ssigma: \partial \Omega \rightarrow \RR^2$ is
an unknown density function. The BIE reformulation has several immediate
advantages including a dimension reduction since the unknown density
function is defined only on $\partial \Omega$, and the far field
boundary condition of the PDE is automatically satisfied. The density
function $\ssigma$ is chosen to satisfy the boundary condition, and this
results in a BIE 
\begin{align*}
  \gg(\xx) = \lambda \ssigma(\xx) + 
    \int_{\partial\Omega} G(\xx - \yy) \ssigma(\yy) \, ds_\yy,
    \quad \xx \in \bd\Omega,
\end{align*}
where $\gg$ is the Dirichlet boundary condition, and $\lambda$ is a
constant that depends on the choice of $G$. Note that if the
well-potential is $f(u) = \tfrac{1}{2}u^2$, then
equation~\eqref{eqn:phase} becomes the so-called linear screened Laplace
equation and can be written as a BIE with an unknown scalar-valued
density function.

In addition to the dimension reduction and far field conditions being
automatically solved, additional advantages of a BIE formulation are
that they are high-order, or even spectrally, accurate; their
discretization results in a well-conditioned linear system that can be
solved iteratively with a mesh-independent number of iterations of a
Krylov method such as the generalized minimal residual method (GMRES);
and complex geometries are easier to address than mesh based methods
such as the finite element method Figure~\ref{fig:fem_vs_bie}. However,
BIEs do have several challenges: nearly touching grains requires a
quadrature rule for nearly-singular integrals; their discretization
results in dense matrices; and they are not immediately suitable to
non-linear PDEs. The BIE community, including PI Quaife, has developed a
variety of tools to address these numerical challenges, and we discuss
our planned approach in Section~\ref{sec:specificaim2}.

%The boundary integral equation (BIE) method is the ideal numerical
%method for the present study in terms of computational cost and
%accuracy. Boundary integral equations require linear PDE and so we take
%$f(\eta) = \tfrac{1}{2}\eta^2$, yielding in () the so called screened
%Laplace equation $-\rho^2 \Delta \eta + \eta$.  This is a Helmholtz
%equation with purely imaginary wave number, enabling us to represent the
%solution in terms of layer potentials:
%\begin{equation}
%  \eta(\xx) = \int_{\partial \Omega} \frac{\partial G}{\partial \nu}
%  \left(\frac{\xx-\yy}{\rho}\right) \sigma(\yy) \,ds_{\yy}
%\end{equation}
%where $\frac{\partial G}{\partial \nu}$ is the normal derivative of the
%fundamental solution of the screened Laplace equation
%(a Bessel function in two space dimensions and
%an exponential times the standard fundamental solution of the Laplace equation
%in three space dimensions).
%This representation allows us to transfer the boundary value problem ()-()
%to an equation only involving in terms of domain boundary.  
%
%The numerical advantages of doing so are as follows.  If $h = N^{-1}$ is
%the characeristic grid spacing of the grain boundaries, then a finite
%element or finite difference discretization of ()-() requires on the
%order of $N_b N^n$ grid points for dense suspensions.  In contrast BIE
%require onely With finite elements, mesh regeneration would be needed at
%each time step since the domain is moving.  In the boundary integral
%approach, collocation points can be selected for the initial grain
%configuration and then used throughout the time-course.  As a downside,
%integral equations require specialized quadrature routines to resolve
%self interactions and nearly touching grain boundaries.  Finally, the
%far field boundary conditions which are important for imposed flow
%problems automatically hold, whereas with finite element and finite
%difference methods artificial boundary conditions must be imposed.
\begin{wrapfigure}[10]{r}{2in}
  \todo[inline]{A figure comparing FEM versus BIE meshes}
  \caption{\label{fig:fem_vs_bie} A finite element method requires
  meshing the entire domain (typically into triangles), while a BIE
  only requires meshing the boundary of the domain.}
\end{wrapfigure}

@@
